---
title: "Getting Started with fmriproj: Direct MVPA from fMRI Time-Series"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with fmriproj}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `fmriproj` package revolutionizes fMRI analysis by enabling MVPA directly from time-series data, bypassing traditional two-stage approaches. This vignette will guide you through:

1. Basic single-subject analysis
2. Searchlight MVPA with rMVPA integration  
3. Optimizing HRF parameters for your data
4. Advanced features and diagnostics

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("bbuchsbaum/fmriproj")
```

```{r setup}
library(fmriproj)
library(fmrireg)  # For event models
library(rMVPA)    # For MVPA analyses
```

## Quick Start: Your First Analysis

Let's start with a complete example that you can adapt to your data:

```{r quickstart, eval=FALSE}
# Load your data
bold_data <- read_nifti("subject01_bold.nii")    # Time x Voxels
events <- read.csv("subject01_events.csv")       # Trial timing
brain_mask <- read_nifti("brain_mask.nii")       # Binary mask

# Create event model (using fmrireg)
event_model <- fmrireg::event_model(
  onsets = events$onset_time,        # When each trial started
  conditions = events$condition,     # Trial labels (e.g., "face", "house")  
  blocks = events$run_number        # Which run each trial belongs to
)

# Run searchlight MVPA with projection
results <- run_searchlight_projected(
  Y = bold_data,
  event_model = event_model,
  radius = 3,                       # Searchlight radius in voxels
  model_name = "sda_notune"         # Classifier (from rMVPA)
)

# Save results
write_nifti(results$accuracy_map, "mvpa_accuracy.nii")
```

That's it! The projection happens automatically behind the scenes.

## Step-by-Step Walkthrough

### Step 1: Understanding Your Event Model

The event model tells fmriproj when trials occurred:

```{r event_model_example}
# Example event structure
events <- data.frame(
  onset_time = c(10, 25, 40, 55, 70, 85),  # Seconds from scan start
  condition = c("A", "B", "A", "B", "A", "B"),
  run = c(1, 1, 1, 1, 1, 1)
)

# Create the model
event_model <- list(
  onsets = events$onset_time,
  conditions = events$condition,
  blocks = events$run,
  n_time = 100  # Total number of time points
)

print(event_model)
```

### Step 2: Direct Projection Pipeline

Here's what happens under the hood, which you can also do manually:

```{r manual_pipeline}
# Simulate some data for demonstration
set.seed(42)
n_time <- 100
n_voxels <- 1000
Y <- matrix(rnorm(n_time * n_voxels), nrow = n_time)

# 1. Build trial-specific design matrix
design <- build_design_matrix(
  event_model = event_model,
  hrf_basis_matrix = NULL  # Uses default HRF
)

cat("Design matrix:", dim(design$X), "\n")
cat("Trials:", length(event_model$onsets), "\n")
cat("HRF bases:", ncol(design$hrf_info$basis), "\n")

# 2. Create projector
projector <- build_projector(
  X_theta = design$X,
  lambda_global = 0.1  # Small ridge regularization
)

# 3. Project a searchlight
searchlight_data <- Y[, 1:50]  # First 50 voxels
proj_result <- adaptive_ridge_projector(
  Y_sl = searchlight_data,
  projector_components = projector,
  lambda_adaptive_method = "EB"  # Empirical Bayes
)

# 4. Collapse HRF components to get trial patterns
trial_patterns <- collapse_beta(
  Z_sl_raw = proj_result$Z_sl_raw,
  N_trials = length(event_model$onsets),
  method = "rss"  # Root-sum-square
)

cat("Trial patterns:", dim(trial_patterns$A_sl), "\n")
```

### Step 3: Integration with rMVPA

The seamless integration with rMVPA makes complex analyses simple:

```{r rmvpa_integration, eval=FALSE}
# Traditional rMVPA (with beta maps)
dataset <- mvpa_dataset(beta_maps, mask)
model <- mvpa_model(load_model("sda_notune"), dataset, design)
results <- run_searchlight(model, radius = 3)

# With fmriproj (direct from time-series)
dataset <- mvpa_dataset(time_series, mask)  # Note: time-series, not betas!
model <- mvpa_model(load_model("sda_notune"), dataset, design)
results <- run_searchlight_projected(model, radius = 3)  # Just add "_projected"
```

## Common Analysis Patterns

### Pattern 1: Whole-Brain Searchlight Classification

```{r whole_brain, eval=FALSE}
# Set up cross-validation
cv_scheme <- blocked_cross_validation(events$run)

# Configure projection
projection_opts <- list(
  lambda_adaptive_method = "EB",     # Adaptive regularization
  collapse_method = "rss",          # How to combine HRF components
  use_progressive_projection = TRUE, # Dimensionality reduction
  pp_dims = 10                      # Reduce to 10 dimensions
)

# Run analysis
results <- mvpa_searchlight(
  Y = bold_data,
  event_model = event_model,
  mask = brain_mask,
  radius = 3,
  classifier = "sda_notune",
  cross_validation = cv_scheme,
  projection_opts = projection_opts
)

# Get accuracy map
accuracy_map <- results$performance_map
```

### Pattern 2: ROI-Based Analysis

```{r roi_analysis, eval=FALSE}
# Extract ROI data
roi_mask <- read_nifti("visual_cortex_mask.nii")
roi_indices <- which(roi_mask > 0)
roi_data <- bold_data[, roi_indices]

# Project entire ROI at once
roi_patterns <- project_trials(
  Y = roi_data,
  event_model = event_model,
  lambda_method = "EB"
)

# Now use standard machine learning
library(caret)
accuracy <- train(
  x = roi_patterns,
  y = as.factor(events$condition),
  method = "svmLinear",
  trControl = trainControl(method = "LOOCV")
)
```

### Pattern 3: Representational Similarity Analysis (RSA)

```{r rsa_analysis, eval=FALSE}
# Project to get trial patterns
trial_patterns <- project_trials(
  Y = roi_data,
  event_model = event_model
)

# Compute neural RDM
neural_rdm <- as.matrix(dist(trial_patterns, method = "correlation"))

# Compare to model RDM
model_rdm <- build_model_rdm(events$condition)
rsa_score <- cor(neural_rdm[lower.tri(neural_rdm)],
                 model_rdm[lower.tri(model_rdm)])

cat("RSA correlation:", rsa_score, "\n")
```

## Advanced Features

### Optimizing HRF Parameters

If your population has non-standard HRF characteristics:

```{r hrf_optimization, eval=FALSE}
# Define flexible HRF basis
flexible_hrf <- function(theta, time_points) {
  peak_time <- theta[1]    # Time to peak
  width <- theta[2]        # Response width
  
  # Generate double-gamma HRF
  hrf <- dgamma(time_points, shape = peak_time/width, scale = width)
  hrf_max <- max(hrf)
  hrf <- hrf / hrf_max
  
  # Add derivative
  hrf_deriv <- c(0, diff(hrf))
  
  cbind(hrf, hrf_deriv)
}

# Optimize for your data
optimal_hrf <- optimize_hrf_mvpa(
  theta_init = c(5, 1),    # Initial: peak=5s, width=1
  Y = bold_data,
  event_model = event_model,
  hrf_basis_func = flexible_hrf,
  metric = "classification_accuracy"
)

cat("Optimal HRF peak:", optimal_hrf$theta_hat[1], "seconds\n")
```

### Diagnostic Visualizations

Understanding what the projection is doing:

```{r diagnostics, eval=FALSE}
# Run with diagnostics enabled
results <- run_searchlight_projected(
  model,
  radius = 3,
  diagnostics = TRUE
)

# Extract diagnostic maps
diag_maps <- explain_projection_results(results)

# Visualize adaptive lambda values
plot_brain(diag_maps$lambda_map, 
           title = "Adaptive Regularization (λ) Map")

# Show HRF collapse weights
plot_brain(diag_maps$w_maps[,1], 
           title = "HRF Amplitude Weights")

# For a specific searchlight
peak_voxel <- which.max(results$accuracy_map)
plot_searchlight_diagnostics(results, voxel = peak_voxel)
```

## Tips for Success

### 1. Data Preparation

```{r data_prep_tips, eval=FALSE}
# Ensure your timing is accurate
print(summary(diff(event_model$onsets)))  # Check trial spacing

# Remove low-frequency drifts (if not done in preprocessing)
Y_detrended <- detrend_bold(Y, method = "polynomial", degree = 3)

# Consider temporal resolution
TR <- 2  # seconds
if (min(diff(event_model$onsets)) < 3 * TR) {
  warning("Trials may be too close for reliable separation")
}
```

### 2. Parameter Selection

```{r param_tips}
# For different noise levels
projection_params <- list(
  high_SNR = list(lambda_method = "none", lambda_floor = 0.01),
  medium_SNR = list(lambda_method = "EB", lambda_floor = 0.1),
  low_SNR = list(lambda_method = "EB", lambda_floor = 1.0)
)

# For different analysis goals  
collapse_params <- list(
  standard = "rss",        # Good default
  maximize_SNR = "pc",     # When signal is weak
  optimize_class = "optim" # When you have good labels
)
```

### 3. Troubleshooting

```{r troubleshooting}
# Check condition number of design matrix
check_design_quality <- function(event_model) {
  design <- build_design_matrix(event_model)
  projector <- build_projector(design$X, diagnostics = TRUE)
  
  cond_num <- attr(projector, "diagnostics")$cond_R
  if (cond_num > 1e6) {
    warning("Design matrix is poorly conditioned. Check trial timing.")
  }
  
  return(cond_num)
}

# Verify projection works on test data
test_projection <- function(Y_test, event_model) {
  tryCatch({
    result <- project_trials(Y_test[, 1:100], event_model)
    cat("Success! Output dimensions:", dim(result), "\n")
  }, error = function(e) {
    cat("Error:", e$message, "\n")
  })
}
```

## Conclusion

The `fmriproj` package makes direct MVPA from time-series both powerful and accessible. Key advantages:

- **Speed**: 10-100x faster than traditional approaches
- **Accuracy**: Adaptive regularization for each searchlight
- **Flexibility**: Optimize HRF for your data
- **Integration**: Works seamlessly with rMVPA

For more examples and advanced usage, see the other vignettes:

- `vignette("optimizing-hrf-parameters")`: Deep dive into HRF optimization
- `vignette("advanced-projections")`: Custom basis functions and constraints  
- `vignette("benchmarking-performance")`: Speed and accuracy comparisons