---
title: "Running rMVPA Analyses on Time-Series Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running rMVPA Analyses on Time-Series Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This article demonstrates the streamlined interface for running multivariate
pattern analyses directly from fMRI time-series data. The high level
functions `run_searchlight()` and `run_regional()` wrap the full
projection pipeline and automatically create all required `rMVPA` objects.

The refactored interface uses `fmridataset` objects to bundle time-series data,
masks, event tables, and sampling frames together, providing a clean and
consistent API for MVPA analyses.

We assume you have already installed **fmriproj**, **fmrireg**, **rMVPA**,
and **fmridataset** from GitHub.

```{r eval=FALSE}
# Install packages
remotes::install_github("bbuchsbaum/fmrireg")
remotes::install_github("bbuchsbaum/rMVPA")
remotes::install_github("bbuchsbaum/fmridataset")
remotes::install_github("bbuchsbaum/fmriproj")
```

```{r setup}
library(fmriproj)
library(fmrireg)
library(rMVPA)
library(fmridataset)
```

## Creating an fmridataset

The first step is to create an `fmridataset` object that bundles your time-series
data, brain mask, event table, and sampling frame:

```{r create-dataset, eval=FALSE}
# Your time-series data: time x voxels matrix
Y <- matrix(rnorm(10000), nrow = 200)

# Brain mask: logical vector or matrix
brain_mask <- rep(TRUE, 50)

# Event table with trial information
events <- data.frame(
  onset = seq(10, 180, by = 20),
  condition = rep(c("A", "B"), each = 5),
  run = rep(c(1, 2), each = 5)
)

# Sampling frame describing data structure
sampling_frame <- data.frame(
  block = c(1, 2),
  blocklens = c(100, 100)  # 100 time points per run
)

# Create fmridataset
fmri_dset <- fmridataset::matrix_dataset(
  Y,
  mask = brain_mask,
  event_table = events,
  sampling_frame = sampling_frame
)
```

## Basic Searchlight Example

With an `fmridataset` object, running a whole-brain searchlight analysis
requires specifying formulas to select the relevant columns from the event table:

```{r searchlight-simple, eval=FALSE}
# Run searchlight analysis
sl_result <- run_searchlight(
  fmri_dset,
  radius = 3,
  y_formula = ~ condition,     # Select 'condition' as labels
  block_formula = ~ run,       # Select 'run' for cross-validation blocks
  classifier = "sda_notune"
)
```

The returned object is an `rMVPA::searchlight_result` containing
accuracy maps and optional diagnostics.

## Running a Regional Analysis

For a single region of interest, the `run_regional()` function works similarly
but requires a region mask:

```{r regional-simple, eval=FALSE}
# Define region of interest (NeuroVol or logical mask)
roi_mask <- brain_mask  # Same as brain mask for this example

roi_result <- run_regional(
  fmri_dset,
  region_mask = roi_mask,
  y_formula = ~ condition,
  block_formula = ~ run,
  classifier = "sda_notune"
)
```

## Customising Projection Options

Both high level functions accept a `projection_opts` list to control
regularisation and HRF collapse methods:

```{r opts-example, eval=FALSE}
# Custom projection options
opts <- list(
  lambda_adaptive_method = "CV",
  collapse_method = "pc",
  lambda_global = 0.2
)

sl_cv <- run_searchlight(
  fmri_dset,
  radius = 3,
  y_formula = ~ condition,
  block_formula = ~ run,
  projection_opts = opts,
  classifier = "sda_notune"
)
```

## Working with Complex Event Tables

The formula-based design allows for flexible selection of labels and blocks
from complex event tables:

```{r complex-events, eval=FALSE}
# Event table with multiple factors
complex_events <- data.frame(
  onset = seq(10, 180, by = 10),
  task = rep(c("memory", "attention"), 9),
  difficulty = rep(c("easy", "hard"), each = 9),
  session = rep(1:3, each = 6),
  run = rep(1:2, 9)
)

# Create dataset with complex event table
complex_dset <- fmridataset::matrix_dataset(
  Y,
  mask = brain_mask,
  event_table = complex_events,
  sampling_frame = sampling_frame
)

# Use interaction terms or specific columns
sl_result <- run_searchlight(
  complex_dset,
  y_formula = ~ task,           # Main effect of task
  block_formula = ~ session,    # Cross-validate across sessions
  classifier = "sda_notune"
)

# Or use interaction of factors
sl_interaction <- run_searchlight(
  complex_dset,
  y_formula = ~ task:difficulty,  # Task × difficulty interaction
  block_formula = ~ run,
  classifier = "sda_notune"
)
```

## Progressive Projection Integration

The projection framework can be combined with rMVPA's feature selection
mechanisms. However, this requires manual construction of the model components:

```{r pp-example, eval=FALSE}
# Manual model construction for advanced features
Y <- fmridataset::get_data(fmri_dset)
mask <- fmridataset::get_mask(fmri_dset)

# Create base rMVPA objects
dataset <- rMVPA::mvpa_dataset(Y, mask)
design <- create_mvpa_design_from_dataset(fmri_dset, ~ condition, ~ run)
cross_validation <- rMVPA::blocked_cross_validation(
  model.frame(~ run, fmri_dset$event_table)[[1]]
)

# Create model with progressive projection feature selector
model <- rMVPA::mvpa_model(
  model = rMVPA::load_model("sda_notune"),
  dataset = dataset,
  design = design,
  crossval = cross_validation
)

# Add progressive projection feature selector
model$feature_selector <- pp_feature_selector(dims = 5)

# Run with custom model
pp_result <- rMVPA::run_searchlight(model, radius = 3)
```

## Interpreting Results and Diagnostics

The `searchlight_result` and `regional_mvpa_result` objects contain
performance values and optional diagnostic information about the
projection step:

```{r results-example, eval=FALSE}
# Access performance results
performance_map <- sl_result$performance

# Plot performance
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot(sl_result)
}

# Access diagnostic information if available
if (!is.null(sl_result$diagnostics)) {
  # Projection-specific diagnostics
  lambda_values <- sl_result$diagnostics$lambda_sl_chosen
  hrf_weights <- sl_result$diagnostics$w_sl
}
```

## Data Format Requirements

The `fmridataset` approach requires specific data formats:

- **Time-series data**: `time × voxels` matrix
- **Event table**: Data frame with `onset` column and factor columns for conditions
- **Sampling frame**: Data frame describing the temporal structure (runs/blocks)
- **Mask**: Logical vector or `NeuroVol` object defining brain regions

```{r data-format, eval=FALSE}
# Validate your dataset
str(fmri_dset)

# Check event table structure
head(fmri_dset$event_table)

# Check sampling frame
print(fmri_dset$sampling_frame)

# Verify time dimensions match
cat("Time points in data:", nrow(fmridataset::get_data(fmri_dset)), "\n")
cat("Total time from sampling:", sum(fmri_dset$sampling_frame$blocklens), "\n")
```

For more details on creating `fmridataset` objects, see the `fmridataset` package documentation.

