---
title: "A Unified Workflow: From fmridataset to MVPA Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Unified Workflow: From fmridataset to MVPA Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette demonstrates the streamlined analysis pipeline using
`fmridataset` objects with `fmriproj`. A single dataset encapsulates the
time-series, mask and event information which is then passed directly to
`run_searchlight()` or `run_regional()`.

## Creating an fmridataset

```{r}
# simulate small data for illustration
Y <- matrix(rnorm(1000), nrow = 100)
mask <- rep(TRUE, 10)
events <- data.frame(
  onset = seq(1, 100, by = 5),
  condition = rep(c("A", "B"), each = 10),
  run = 1
)

library(fmridataset)
dset <- matrix_dataset(
  Y, mask = mask,
  event_table = events,
  sampling_frame = data.frame(block = 1, blocklens = nrow(Y))
)
```

## Running a Searchlight Analysis

```{r eval=FALSE}
sl_res <- fmriproj::run_searchlight(
  dset,
  radius = 2,
  y_formula = ~condition,
  block_formula = ~run
)
```

## Running a Regional Analysis

```{r eval=FALSE}
roi_res <- fmriproj::run_regional(
  dset,
  region_mask = mask,
  y_formula = ~condition,
  block_formula = ~run
)
```

Both functions automatically perform trial projection and interface
with `rMVPA` to produce MVPA results.
